Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-09-07 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd

<!-- Tipping points and the frequency of a lack of data throughout state-space -->

@@

add references to [@Fischer2009, @Polasky2011] as persective on combining complex / critical dynamics + control theory

@@

Decision-theoretic (or optimal control) tools require a model that can assign probabilities of
future states (e.g. stock size of a fishery) given the current state and a
proposed action (e.g. fishing harvest or effort).  The decision maker then
seeks to determining the course of actions (also referred to as the policy)
that maximizes the expected value of some objective function (such as
net present value derived from the resource over time.  (The approach
can be adapted to permit alternatives to maximizing expectation, such
as minimizing the maximum cost or damage that might be incurred. See
@Polasky2011).
Management frequently faces a sequential decision-making problem -- after 
selecting an action, the decision-maker may receive new information about 
the current state and must again choose an appropriate action -- i.e. setting
the harvest limits each year based on stock assessments the year prior.

More historical context, e.g. [@Reed1979, @Mangel1985]

#### Parametric models

Traditional approaches to optimal control (Pontryagin's principle, stochastic
dynamic programming) rely on knowledge of the state equation.  Such simple 
parametric models, frequently justified by mechanistic underpinnings are used
to provide probabilities of future states.  Fisheries literature, which enjoys
some of the longest history in the use of the optimal control framework, provides
a clear example [ ] . Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
Parametric models can reflect variability introduced by a stochastic environment
or demographic process my incorporating random shocks from a given distribution.

>>>>

<!-- More on complex dynamics and not having the correct model

Cite [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
etc

also older stuff on model uncertainty

<!-- More on the lack of data throughout the relevant state-space
     and how any concern about potential tipping points indicates 
     that the data exhibits this bias / problem.

@@

@@

Decision-theoretic or optimal control tools require a model that can assign probabilities of
future states (e.g. stock size of a fishery) given the current state and a
proposed action (e.g. fishing harvest or effort).  
Management frequently faces a sequential decision-making problem -- after 
selecting an action, the decision-maker may receive new information about 
the current state and must again choose an appropriate action -- such as setting
the harvest limits each year based on stock assessments the year prior.

The decision maker typically seeks to determining the course of actions (also referred to as the policy)
that maximizes the expected value of some objective function such as
net present value derived from the resource over time.  
Though much can be said on how to choose this value function appropriately 
(e.g. see [@Halpern2013](http://doi.org/10.1073/pnas.1217689110 
"Achieving the triple bottom line in the face of inherent trade-offs 
among social equity, economic return, and conservation.")) we will
assume this is given.  (Nor is this approach necessarily constrained to 
maximizing the expectated value of such a function - the decision-theoretic
framework can be adapted to alternatives such as minimizing the maximum 
cost or damage that might be incurred; see @Polasky2011).

In representing future states with probabilities and maximizing expectations,
this approach provides a natural framework for handling uncertainty.

The value function typically depends on the action or policy taken, as well as
the state of the system, in each interval of time. The state of the system,
in turn, is usually described by a dynamical model.
